---
title: "Untitled"
author: "David Contreras Carretero"
date: "17/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"

### Excecutive Summary

### Exploratory Analysis


```{r cars}
head(mtcars)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
```


The boxplot shows how 

### Is an automatic or manual transmission better for MPG?


Observing the box plot (fig. 1 Appendix 1) we it seems that in mort that 75% of the cases automatic vehicles are less efficient than those with manual transmission.



### Appendix 1



```{r }
boxplot(mpg~am, data = mtcars,
        xlab = "Transmission (Automatic/Manual)",
        ylab = "Miles per Gallon",
        main = "Miles per galon by transmission type")
mtext("Figure 1",side=1,line=4)
```

