---
title: "Connections"
author: "DCC"
date: "5/6/2020"
output: html_document
---

# CONNECTIONS

* *file*, connection to a file
* *gzfile*, connection to a gzip file
* *bzfile*, bzip2
* *url*, connection to webpage

## File Connections

```{r file}
str(file)
```

* *description*, name of the file
* *open*, indicate mode
  + *"r"* read only
  + *"w"* write, new file
  + *"a"*, append
  + *"rb"*, *"wb"*, *"ab"* reading, writing, or appending in binary mode "windows"
  
## *con*

```{r con file}
con <- file("../datos/oleajePuertos.csv", "r")
data <- read.csv(con)
close(con)
data
```

also:

``` {r read csv} 
data <- read.csv("../datos/oleajePuertos.csv", sep = ";")
data
```

##Reading Lines of a text file

using *con* to read n lines from a compressed file

```{r con gzip}
con <- gzfile("../datos/01001bsc.csv.gz")
x <- readLines(con, 10)
x
```

### read lines from webpage

```{r read webpage}
## this might take some time
con <- url("http://www.ororo.tv", "r")
x <- readLines(con)
head(x)
```


