---
title: "Control Structures"
author: "DC"
date: "8/6/2020"
output: html_document
---

# CONTROL STRUCTURES

##   *if - else*


if (condition){
  ## some code
} else {
  ## some more
}

if (condition) {
  
} else if (condition 2){
  
  
}  else {
    
 }


Special R if-else construction:

``` {r}
x <- 6
y <- if (x >3) {
  10
}else  {
  0
}
```


## *For loops*


```{r 4 loops}
x <- c("a", "b", "c", "d")

for (i in 1:4) {
   print(x[i])
}

for (i in seq_along(x)) { ## creates integer sequence equal to the length                            ## of x (4)
  print(x[i])
}

for (letter in x) {
  print(letter) 
}

for (i in 1:4) print (x[i])
```


## Nested for loops

``` {r nested loops}
x <- matrix(1:6,2,3)

for (i in seq_len(nrow(x))) {
    for (j in seq_len(ncol(x))) {
        print(x[i, j])
    }
  
}
```

## While loops

``` {r}
count <- 0
while(count < 10){
    print(count)
    count <- count + 1
}
````


``` {r while z moves depending on random coin throw }
z <- 5

while (z >= 3 && z <= 10) {
  print (z)
  coin <- rbinom(1, 1, 0.5)
 
  if(coin == 1 ){  ## random walk
      z <- z + 1
  } else {
      z <- z - 1
  }
}
```


## Repeat, Next, Break


#*Repeat* 
Dangerous because there's no guarantee it wil stop. Better to set a hard limit on the number of iterations.


*x <- 1*
*tol <- 1e-8*

*repeat {*
  *x1 <- computeEstimate()*   
  
  *if(abs(x1 - x0) < tol ) {*
    *break*
  *} else {*
    *x0 <- x1*
  *}*
*}*

#*next, return*

*next* skips iteration

``` {r}
for (i in 1:100){
    if (i <= 20) {
        ## Skip the fist 20 iterations
        next
    }
  ##do something here 
}
  
```
  
*return*, signals that a functin should exit and return a value.



