---
title: "DATAE AND TIMES"
output: html_notebook
---


* Dates are represented by the _Date_ class
* Times ar represnted by the _POSIXct_ or or th _POSIXlt_ class
* Dates ar stored internally as the __number of days since 01/01/1970__
* And times as the __number of seconds__ since the same date


## Dates in R

``` {r}
x <- as.Date("1970-01-01")
x
unclass(x)   ## 0 days from 1970-01-01
x <- as.Date("1970-01-02")
unclass(x)   ## 1 days from 1970-01-01
```


## Times in R

### _POSIXct_ , _POSIXlt_

*___POSIXct___ a very large integer, it use a useful class whe you want store times in a data frame
* ___POSIXlt___  its a list that stores other useful information like the day of the week, day of the year, month, day of the month


There's some generic functions that w9ork on dates an times : _weekdays_, _months_, _quarters_...  Those functions operate on objects of class _POSIXct_, _POSIClt_ or _Date_

```{r}
weekdays(x)
months(x)
quarters(x)
```

Times can be coerced from a character string uising the as.POSIXlt or as.POSIXct function.

```{r}
x <- Sys.time()
x
p <- as.POSIXlt(x)
names(unclass(p))
p$sec
```

_POSICXct_ does not allow to extract elements because is not a list, but can be coercet to _POSICXlt_ as above.






