---
title: "Functions"
author: "DC"
date: "9/6/2020"
output: html_document
---

#  FUNCTIONS

Functios in R are objects of the class "function" and ar created with *function()* directive

That means that they can be treated as any other R objects.

* Functions can be passed as arguments to other funcitons
* Functions can be nested. Define one function inside another.


``` {r}
f <- function( arguments) {
  ## actions
}
```

Fuctions return the value of the last expression evaluated, there's no need of a return statment, tot i que existeix una funció return()

## Function arguments
Named arguments can have default values

* Fromal arguments, those included in definition
* *formals* function returns a list of the formal arguments of a function
* Not every function cal in R makes use of all formal arguments
* Function arguments can be missing or might have default values

## Argument Matching

Can be matchet by position or by name.

```{r equivalent expression }
mydata <- rnorm(100)
sd(mydata)
sd ( x = mydata)
sd(x = mydata, na.rm = FALSE)
sd(na.rm = FALSE, x = mydata)
sd(na.rm = FALSE, mydata)
```

Position and name matching can be mixed.

```{r args}
args(lm)
```


*lm(data = mydata, y - x, model = FALSE, 1:100)*
*lm(y - x, mydata, 1:100, model = FALSE)*

You can also partially match the name of the argument. Mostly for interactive work in console, not for programing

The order is exact match -> partial match -> positional

## The "..." Argument

``` {r }
args(paste)
args(cat)
```

 ... the number or arguments passed to the function cannot be known in advance.
 
## Arguments after the "..." argument

Arguments after ... on the argument list must be named explicitly and cannot be partially matched.

``` {r}
args(paste)
paste("a","b", sep = ":")
paste("a","b", se = ":")  ## "se"  does not partially match to sep
```




