---
title: "SCOPING RULES"
output: html_notebook
---


## Symbol Binding

```{r}
search()
```

*search()* displays the search list R would use to bind a value to a symbol.

It is possible to overwrite a existing function within the immediate enviroment. Consists in all the R packages loaded into R ( your workspace).

* __*Global environment*__ , is the user's workspace, is __always the first element__ of the search list and __*base*__ package is __always__ the __last__
* The __order__ of packages on the search list __matters__.
* Packages loaded on startup can be configurated.
* Loading a __package__ through __*library* sets its namespace in position 2__ of the search list.
* R has separate namespaces for functions and non-funcions, __it's possible__ to have an __object and a function with the same name__.


## Scoping Rules

* __Scoping rules__ determine how a value is associeted with a free varialbe in a function
* __Lexical or static scoping__
* Lexical scopint applies perticularly well to __statistical computations simplification__


In the next function scope rules determine the value to be assingned to "z" _free variable_
```{r} 
f <- function (x, y) {
   x^2 + y / z
}
```


### Lexical Scoping

_the values of free variables are searched for in the __environment__ in which the function was defined._

* ___environment___, a collenction of (symbol, value) pairs
* Every enviroment has a parent evironment; and could have 0 or many children
* The only environment without a parent is the empty environment
* A function + an environment is a __closure__ or __function closure__

 Search order:
  1. Environment in which the function was created
  2. Parent environment
  3. Starts going down the search list ( global env, tools:rstudio, package:...)
  4. Reach __empty environment__ without matching then error is thrown


* Typically, a function is defined in a global enviroment, so the values of free variables are just found in the user's workspace
* In R is __possible to define functions inside other functions__, and that function can be returned from the function where it was defined, so that function's __enviroment would be the insides of the other function__
 (usually called constructor function)
 
 
Constructor function example:

``` {r}
make.power <- function(n){  ##constructor function power
      pow <- function(x) {
         x^n
      } 
      pow
}

cube <- make.power(3)   ##construction of function cube
square <-make.power(2)  ##construction of function square
cube(3)
square(3)
```


  _environment_ function, allows to see what's in the environment of a function.
  
```{r environment and get}
ls(environment(cube))
get("n",environment(cube))

ls(environment(square))
get("n",environment(square))
```


##  Lexical vs. Dynamic Scoping

Dynamic scoping is common in most programming languages as C

``` {r}
y <- 10

f <- function(x) {
    y <- 2
    y^2 + g(x)
}

g <- function(x) {
    x*y
  
}

f(3)
```

* Lexical scoping looks up for the value of _y_ in the function _g_ in the __environment in which the function was defined__, in this case the global environment, the value is _10_

* Dynamic scoping, looks up for the value of _y_ in the __environment from which the function was called__ (_calling environment_), it woud be 2

* If function is defined in the global environment and called from teh global environment, __the defining and the calling environment coud be the same__

* Other lenguages with lexical scoping are:
    + Scheme
    + Perl
    · Python
    + Common Lisp (all languages converge to Lisp)
    
    
## Optimization Example

   _optim_, _nlm_, _optimize_
   
### Maximizing a Normal Likelyhood

``` {r}
make.NegLogLik <- function(data, fixed=c(FALSE,FALSE)){
    params <- fixed
    function(p) {
       params[!fixed] <- p
       mu <- params[1]
       sigma <- params[2]
       a <- -0.5*length(data) * log (2*pi*sigma^2)
       b <- -0.5*sum((data-mu)^2) / (sigma^2)
       -(a + b)
    }

}
  
set.seed(1); normals <- rnorm(100, 1, 2)
nLL <- make.NegLogLik(normals)
nLL
ls(environment(nLL))
```

















