---
title: "Subsetting"
author: "DCC"
date: "5/6/2020"
output: html_document
---

# SUBSETTINGS

## Basics

Operators to extract subsets of R objects

* *[* returns an object of the same class as the original
* *[[* extract elements of a list or data frame, it can only be used to extract a single element, the class will not necessarily be a list o data frame
* *$* extract elements of a list or data frame by name

```{r subsetting}
x <- c("a","b","c","c","d","a")
x[1]
x[2]
x[1:4]
x[x>"a"]
u <-x > "a"
u
x[u]
```
## Subsetting Lists

```{r subsetting lists}
x <- list (foo = 1:4, bar = 0.6)
x[1]    ## retruns a list
x[[1]]  
x$bar
x[["bar"]]
x["bar"]  ## retruns a list
```

To extract multiple elements from a list use only *[* 

```{r multiple elements}
x <- list(foo = 1:4, bar = 0.6, baz = "fellow")
x[c(1,3)]
```


Difference between *$* and *[[* is that *[[* allows to use a variable as element name, that allows to calculate dynamically the element to extract.

```{r doublebracket vs dollarsign}
x <- list (foo = 1:4, bar = 0.6, baz = "yellow")
name <- "foo"
x[[name]] ## computed name for foo
x$name
x$foo
```

### Subsetting Nested Elements of a List

*[[* can take an integer sequence.

``` {r nested}
x <- list ( a= list(10,20,14), b = c(3.14,5.23))
x[[c(1,3)]]
x[[1]][[3]]
x[[c(2,1)]]
```

## Subseting Matrices

Usual way, *(i,j)* indices

``` {r}
x <- matrix(1:6, 2,3)
x
x[1,2]
x[2,1]
x[1,]
x[,1]
```

With matrices *[* does not returns a matrix but a vector of whatever length, that feature can be disabled by using the paramenter *drop = FALSE*, then it will return a 1 x n matrix

``` {r}
x <- matrix(1:6, 2, 3)
x[1, 2] ## vector
x[1, 2, drop = FALSE] ## matrix
x[1, , drop = FALSE] ## matrix
```

## Partial Matching

Partial matching of names is allowed with *[[* and *$*

``` {r}
x <- list(aadvark = 1:5)
x$a
x[["a"]]   ## does not allows partial matching by default
x[["a", exact = FALSE]]
```

## Subseting  - Removin NA values

```{r}
x <- c(1,2,NA,4,NA,5)
bad <- is.na(x)  ## logical vector indicating wherethere the element is NA
x[!bad]        ## print negation of bad is good
```

Multiple vectors with *complete.cases()*

```{r}
x <- c(1,2,NA,4,NA,5)
y <- c("a","b",NA,"d",NA,"F")
good <- complete.cases(x,y)
good
x[good]
y[good]
```


``` {r}
df <- data.frame(name = c("Jon", "Bill", "Maria",NA, "Foo", "Bar","Norf"),
                 age = c(23, 41, 32,12, 23, NA, 67)
                 )
df
good <- complete.cases(df)
df[good,]


